Artin Group Presentations Arising from Cluster Algebras

### 2014-08-01 Paper.tex
"edits to section 4, corrected t(a,b)

In this section, we shall review the structure of diagrams of finite type, and how their cycles are affected by mutation. This section is simply a recap of \cite[Section 2]{BM13}. First, in Proposition ~\ref{cycle_types}, all types of chordless cycles in diagrams of finite type are classified. Second, in Corollary ~\ref{local_three_picture} all possible local pictures between a mutated vertex and two adjacent vertices are drawn. Finally, in ~\ref{lem:chordless_cycles}, all chordless cycles introduced from a mutation are drawn. These three lemmas will be crucial in proving the main result Theorem ~\ref{thm:main}, as they will allow us to inspect precisely which relations are added and removed after mutating at a prescribed vertex.

@@

\begin{prop}
\label{cycle_types}
\cite[Proposition 2.1]{BM13} Let $\Gamma$ be a diagram of finite type. Then, a chordless cycle in the unoriented graph of $\Gamma$ is cyclically oriented in $\Gamma$. Furthermore, the unoriented graph underlying the cycle must either be a cycle such that all edges have weight 1, a square with two opposite edges of weight 2 and two opposite edges of weight 1, as pictured below, or a triangle with two edges of weight 2 and one of weight 1.
\\
\hspace*{-1.2cm}
\begin{tikzpicture}[baseline=-0.5ex]
 \matrix[matrix of math nodes,column sep={40pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (0,0)
  {
    & &|[name = ul]|\circ& |[name = ur]|\circ&\\
    %
    &|[name=lu]| \circ &&&|[name=ru]| \circ \\
     & |[name = ld]|\circ&&& |[name = rd]|\circ\\
    %
    &&|[name=dl]| \circ &|[name=dr]| \circ &\\
  };
\draw
  (ul) -- node[above] {1} (ur)
  ;
  \draw
  (ul) -- node[left]{1}(lu)
  ;
  \draw
  (dr) -- node[right]{1}(rd)
  ;
  \draw
  (lu) -- node[left]{1}(ld)
  ;
  \draw
  (ru) -- node[right]{1}(rd)
  ;
  \draw
  (ur) -- node[right]{1}(ru)
  ;
  \draw
  (dl) -- node[left]{1}(ld)
  ;
  \draw[dotted]
  (dl) -- (dr)
  ;
 \matrix[matrix of math nodes,column sep={40pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (5,0)
  {
    & |[name = ul]|\circ& |[name = ur]|\circ\\
    %
    &|[name=dl]| \circ &|[name=dr]| \circ \\
  };
\draw
  (ul) -- node[above] {1} (ur)
  ;
  \draw
  (dl) -- node[left]{2}(ul)
  ;
  \draw
  (dr) -- node[below]{1}(dl)
  ;
  \draw
  (ur) -- node[right]{2} (dr)
  ;
   \matrix[matrix of math nodes,column sep={25pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (9,0)
  {
    && |[name = u]|\circ& \\
    %
    &|[name=dl]| \circ & &|[name=dr]| \circ \\
  };
\draw
  (u) -- node[right] {2} (dr)
  ;
  \draw
  (dl) -- node[left]{2}(u)
  ;
  \draw
  (dr) -- node[below]{1}(dl)
  ;

@@

\begin{note}
Let
\begin{align*}
\langle x_i,x_j \rangle ^k = \begin{cases}
(x_ix_j)^{\frac{k}{2}}, &\text{ if }k \equiv 0 \pmod 2\\
(x_ix_j)^{\frac{k-1}{2}}x_i &\text{ if } k \equiv 1 \pmod 2
\end{cases}
\end{align*}
That is, $\langle x_i,x_j \rangle$ is just an alternating sequence of $x_i$ and $x_j$ of length $k$.  We also write $\langle x_i,x_j\rangle^{-k}$ to denote $\left(\langle x_i,x_j\rangle^k\right)^{-1}$

\end{note}
\begin{defn}
\cite[Beginning of section 1.2]{C06}
For $M \in M_{n \times n}(\mathbb Z \cup \infty)$ a matrix whose entries can take values in the real numbers or infinity, satisfying $M_{i,j} = M_{j,i}$ we define the associated {\it Artin group} in terms of generators and relations by
\begin{align*}
A = \langle x_1,\ldots, x_n| \langle x_i,x_j \rangle^{M_{i,j}}\forall i,j \text{ with } m_{i,j}<\infty \rangle,
\end{align*}
\end{defn}

\begin{rem}
Each Artin group has an associated Coxeter group defined by adding in the additional relations $s_i^2 = 1$ for all $i.$ An Artin group is said to be of {\it finite type} if its associated Coxeter group is of finite type. To each Artin group of finite type we can assign it the same Dynkin diagram which is assigned to the Coxeter group associated to the Artin group.
\end{rem}

@@

\begin{note}
Let $(i_0,\ldots, i_{d-1})$ be an ordered tuple such that the subgraph of $\Gamma$ on the vertices $i_0,\ldots, i_{d-1}$ is a chordless cycle, with edges of nonzero weight from $i_k$ to $i_{k+1}$, where subscripts are taken $\pmod d.$ Then, denote $$p(i_a,i_{a+1}) = s_{i_{a+1}}^{-1}s_{i_{a+2}}^{-1}\dots s_{i_{a-2}}^{-1}s_{i_{a-1}}s_{i_{a-2}}s_{i_{a-3}}\dots s_{i_{a+1}} s_{i_{a}}^{-1}.$$ Additionally, let $$t(i_a,i_{a+1}) = s_{i_a} p(i_a,i_{a+1}) p(i_a,i_{a+1})^{-1} s_{i_a}^{-1}.$$
\end{note}

@@

\item[(T3)] Let $(i_0,i_1,\ldots,i_{d-1})$ be an ordered tuple such that the subgraph of $\Gamma$ on the vertices $(i_0,\ldots, i_{d-1})$ is a chordless cycle, with edges of nonzero weight from $i_k$ to $i_{k+1},$ where subscripts $k$ are taken $\pmod d.$ If $(i_0,\ldots, i_{d-1})$ is such an ordered tuple, and additionally, one of the following two conditions hold,
\begin{enumerate}
\item All edges in the chordless cycle are of weight 1 or 2 and the edge $i_{d-1}\rightarrow i_0$ has weight 2.
\item All edges in the chordless cycle have weight $1.$
\end{enumerate}
then, we include the relation
$t(i_0,i_1) = e.$
\end{itemize}
\end{defn}

\begin{rem}
In the above definition, the tuple cycle is ordered, and so we may have multiple relations corresponding to the same set $\{i_0,\ldots, i_{d-1}\},$ corresponding to cyclic rotations of the tuple $(i_0,\ldots, i_{d-1}).$ However, we shall see in Section ~\ref{sec:one_relation} that many of these relations are redundant.
\end{rem}

\begin{rem}
We purposely include relations $(T2),(T3)$ but not $(T1)$ in order to make our relation labeling analogous to that of \cite{BM13} at the beginning of Section 3. Note that if we add the additional relation $(T1)$ as defined at the beginning of Section 3 of \cite{BM13} (namely, if we add $s_i^2 = e$ for all vertices $i$ in $\Gamma$), then we will precisely obtain the group $W_\Gamma$ as defined at the beginning of rem 3 in \cite{BM13}.
\end{rem}

@@

\begin{ex}
The relations $(T2),(T3)$ in that $\Gamma$ is a square with all edges of weight 1 are as follows:
\[\Gamma =\begin{tikzpicture}[baseline=-0.5ex]
\matrix[matrix of math nodes,column sep={40pt,between origins},row
sep={40pt,between origins},nodes={asymmetrical rectangle}] at (6,0)
{
& |[name = ul]|\circ& |[name = ur]|\circ\\
%
&|[name=dl]| \circ &|[name=dr]| \circ \\
};
\draw
(ul) -- node[above] {1} (ur)
;
\draw
(dl) -- node[left]{1}(ul)
;
\draw
(dr) -- node[below]{1}(dl)
;
\draw
(ur) -- node[right]{1} (dr)
;
\node [yshift = .4 cm,xshift = -.4 cm] at (ul){1};
\node [yshift = .4 cm,xshift = .4 cm] at (ur){2};
\node [yshift = -.4 cm,xshift = .4 cm] at (dr){3};
\node [yshift = -.4 cm,xshift = -.4 cm] at (dl){4};
\end{tikzpicture}\]
   \begin{enumerate}
\item[(T2)]
\begin{itemize}
\item $\langle s_1,s_2 \rangle^3 = \langle s_2,s_21 \rangle^3$
\item $\langle s_2,s_3 \rangle^3 = \langle s_3,s_2 \rangle^3$
\item $\langle s_3,s_4 \rangle^3 = \langle s_4,s_3 \rangle^3$
\item $\langle s_4,s_1 \rangle^3 = \langle s_1,s_4 \rangle^3$
\end{itemize}
\item[(T3)]
\begin{itemize}
\item $s_{1}s_{2}^{-1}s_{3}^{-1}s_4s_{3}s_2s_1^{-1}s_2^{-1}s_3^{-1}s_4^{-1}s_3s_2 = e$
\item $s_{2}s_{3}^{-1}s_{4}^{-1}s_1s_{4}s_3s_2^{-1}s_3^{-1}s_4^{-1}s_1^{-1}s_4s_3 = e$
\item $s_{3}s_{4}^{-1}s_{1}^{-1}s_2s_{1}s_4s_3^{-1}s_4^{-1}s_1^{-1}s_2^{-1}s_1s_4 = e$
\item $s_{4}s_{1}^{-1}s_{2}^{-1}s_3s_{2}s_1s_4^{-1}s_1^{-1}s_2^{-1}s_3^{-1}s_2s_1 = e$
\end{itemize}
\end{enumerate}
\end{ex}

\begin{rem}
Note that if $\Gamma$ is the graph associated to a Dynkin diagram, then $W_\Gamma$ as we have defined it is precisely the corresponding Artin group corresponding to that Dynkin diagram. This occurs because, in this case, we have no cycles in $\Gamma,$ and so we only have relation of the form $(T2'),$ which define the Artin group.
\end{rem}

>>>>

In this section, we shall review the structure of diagrams of finite type, and how their cycles are affected by mutation. This section is simply a recap of \cite[Section 2]{BM13}. First, in Proposition ~\ref{cycle_types}, all types of chordless cycles in diagrams of finite type are classified. Second, in Corollary ~\ref{local_three_picture} all possible local pictures between a mutated vertex and two adjacent vertices are drawn. Finally, in Lemma ~\ref{lem:chordless_cycles}, all chordless cycles introduced from a mutation are drawn. These three lemmas will be crucial in proving the main result Theorem ~\ref{thm:main}, as they will allow us to inspect precisely which relations are added and removed after mutating at a prescribed vertex.

@@

\begin{prop}
\label{cycle_types}
\cite[Proposition 2.1]{BM13} Let $\Gamma$ be a diagram of finite type. Then, a chordless cycle in the unoriented graph of $\Gamma$ is cyclically oriented in $\Gamma$. Furthermore, the unoriented graph underlying the cycle must either be a cycle such that all edges have weight 1, a square with two opposite edges of weight 2 and two opposite edges of weight 1, or a triangle with two edges of weight 2 and one of weight 1, as pictured below.
\\
\hspace*{-1.2cm}
\begin{tikzpicture}[baseline=-0.5ex]
 \matrix[matrix of math nodes,column sep={40pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (0,0)
  {
    & &|[name = ul]|\circ& |[name = ur]|\circ&\\
    %
    &|[name=lu]| \circ &&&|[name=ru]| \circ \\
     & |[name = ld]|\circ&&& |[name = rd]|\circ\\
    %
    &&|[name=dl]| \circ &|[name=dr]| \circ &\\
  };
\draw
  (ul) -- node[above] {1} (ur)
  ;
  \draw
  (ul) -- node[left]{1}(lu)
  ;
  \draw
  (dr) -- node[right]{1}(rd)
  ;
  \draw
  (lu) -- node[left]{1}(ld)
  ;
  \draw
  (ru) -- node[right]{1}(rd)
  ;
  \draw
  (ur) -- node[right]{1}(ru)
  ;
  \draw
  (dl) -- node[left]{1}(ld)
  ;
  \draw[dotted]
  (dl) -- (dr)
  ;
 \matrix[matrix of math nodes,column sep={40pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (5,0)
  {
    & |[name = ul]|\circ& |[name = ur]|\circ\\
    %
    &|[name=dl]| \circ &|[name=dr]| \circ \\
  };
\draw
  (ul) -- node[above] {1} (ur)
  ;
  \draw
  (dl) -- node[left]{2}(ul)
  ;
  \draw
  (dr) -- node[below]{1}(dl)
  ;
  \draw
  (ur) -- node[right]{2} (dr)
  ;
   \matrix[matrix of math nodes,column sep={25pt,between origins},row
    sep={40pt,between origins},nodes={asymmetrical rectangle}] at (9,0)
  {
    && |[name = u]|\circ& \\
    %
    &|[name=dl]| \circ & &|[name=dr]| \circ \\
  };
\draw
  (u) -- node[right] {2} (dr)
  ;
  \draw
  (dl) -- node[left]{2}(u)
  ;
  \draw
  (dr) -- node[below]{1}(dl)
  ;

@@

\begin{note}
Let
\begin{align*}
\langle x_i,x_j \rangle ^k = \begin{cases}
(x_ix_j)^{\frac{k}{2}}, &\text{ if }k \equiv 0 \pmod 2\\
(x_ix_j)^{\frac{k-1}{2}}x_i &\text{ if } k \equiv 1 \pmod 2
\end{cases}
\end{align*}
That is, $\langle x_i,x_j \rangle$ is just an alternating sequence of $x_i$ and $x_j$ of length $k$.  We also write $\langle x_i,x_j\rangle^{-k}$ to denote $\left(\langle x_i,x_j\rangle^k\right)^{-1}$
\end{note}

\begin{defn}
\cite[Beginning of section 1.2]{C06}
For $M \in M_{n \times n}(\mathbb Z \cup \infty)$ a matrix whose entries can take values in the real numbers or infinity, satisfying $M_{i,j} = M_{j,i},$ we define the associated {\it Artin group} in terms of generators and relations by
\begin{align*}
A = \langle x_1,\ldots, x_n| \langle x_i,x_j \rangle^{M_{i,j}}\forall i,j \text{ with } M_{i,j}<\infty \rangle,
\end{align*}
\end{defn}

\begin{rem}
Each Artin group has an associated Coxeter group defined by adding in the additional relations $s_i^2 = e$ for all $i.$ An Artin group is said to be of {\it finite type} if its associated Coxeter group is of finite type. To each Artin group of finite type we can assign to it the same Dynkin diagram which is assigned to the Coxeter group associated to the Artin group.
\end{rem}

@@

\begin{defn}
\label{defn:cycle_tuple}
Let $(i_0,\ldots, i_{d-1})$ be an ordered tuple such that the subgraph of $\Gamma$ on the vertices $i_0,\ldots, i_{d-1}$ is a chordless cycle, with edges of nonzero weight from $i_k$ to $i_{k+1}$, where subscripts are taken $\pmod d.$ Call such an ordered tuple a {\it chordless cycle tuple.} Then, denote $$p(i_a,i_{a+1}) = s_{i_{a+1}}^{-1}s_{i_{a+2}}^{-1}\dots s_{i_{a-2}}^{-1}s_{i_{a-1}}s_{i_{a-2}}s_{i_{a-3}}\dots s_{i_{a+1}}.$$ Additionally, let $$t(i_a,i_{a+1}) = [s_{i_a},p(i_a,i_{a+1})]$$ where $[a,b] = aba^{-1}b^{-1}$ is the commutator.
\end{defn}

@@

\item[(T3)] Let $(i_0,i_1,\ldots,i_{d-1})$ be a chordless cycle tuple, as defined in Definition ~\ref{defn:cycle_tuple}. If additionally one of the following two conditions hold,
\begin{enumerate}
\item All edges in the chordless cycle are of weight 1 or 2 and the edge $i_{d-1}\rightarrow i_0$ has weight 2,
\item All edges in the chordless cycle have weight $1,$
\end{enumerate}
then, we include the relation
$t(i_0,i_1)=e.$
\end{itemize}
\end{defn}

\begin{rem}
In the above definition, the chordless cycle tuple is ordered, and so we may have other relations corresponding to chordless cycle tuples which are cyclic reorderings the chordless cycle tuple $(i_0,\ldots, i_{d-1}).$ However, we shall see in Section ~\ref{sec:one_relation} that many of these relations are redundant.
\end{rem}

\begin{rem}
We purposely include relations $(T2),(T3)$ but not $(T1)$ in order to make our relation labeling analogous to that of \cite{BM13} at the beginning of Section 3. Note that if we add the additional relation $(R1)$ as defined at the beginning of Section 3 of \cite{BM13} (namely, if we add $s_i^2 = e$ for all vertices $i$ in $\Gamma$), then we will precisely obtain the group $W_\Gamma$ as defined at the beginning of in \cite[Remark 3]{BM13}.
\end{rem}

@@

\begin{ex}
The relations $(T2),(T3)$ in that $\Gamma$ is a square with all edges of weight 1 are as follows:
\[\Gamma =\begin{tikzpicture}[baseline=-0.5ex]
\matrix[matrix of math nodes,column sep={40pt,between origins},row
sep={40pt,between origins},nodes={asymmetrical rectangle}] at (6,0)
{
& |[name = ul]|\circ& |[name = ur]|\circ\\
%
&|[name=dl]| \circ &|[name=dr]| \circ \\
};
\draw
(ul) -- node[above] {1} (ur)
;
\draw
(dl) -- node[left]{1}(ul)
;
\draw
(dr) -- node[below]{1}(dl)
;
\draw
(ur) -- node[right]{1} (dr)
;
\node [yshift = .4 cm,xshift = -.4 cm] at (ul){1};
\node [yshift = .4 cm,xshift = .4 cm] at (ur){2};
\node [yshift = -.4 cm,xshift = .4 cm] at (dr){3};
\node [yshift = -.4 cm,xshift = -.4 cm] at (dl){4};
\end{tikzpicture}\]
   \begin{enumerate}
\item[(T2)]
\begin{itemize}
\item $\langle s_1,s_2 \rangle^3 = \langle s_2,s_1 \rangle^3$
\item $\langle s_2,s_3 \rangle^3 = \langle s_3,s_2 \rangle^3$
\item $\langle s_3,s_4 \rangle^3 = \langle s_4,s_3 \rangle^3$
\item $\langle s_4,s_1 \rangle^3 = \langle s_1,s_4 \rangle^3$
\end{itemize}
\item[(T3)]
\begin{itemize}
\item $s_{1}s_{2}^{-1}s_{3}^{-1}s_4s_{3}s_2s_1^{-1}s_2^{-1}s_3^{-1}s_4^{-1}s_3s_2 = e$
\item $s_{2}s_{3}^{-1}s_{4}^{-1}s_1s_{4}s_3s_2^{-1}s_3^{-1}s_4^{-1}s_1^{-1}s_4s_3 = e$
\item $s_{3}s_{4}^{-1}s_{1}^{-1}s_2s_{1}s_4s_3^{-1}s_4^{-1}s_1^{-1}s_2^{-1}s_1s_4 = e$
\item $s_{4}s_{1}^{-1}s_{2}^{-1}s_3s_{2}s_1s_4^{-1}s_1^{-1}s_2^{-1}s_3^{-1}s_2s_1 = e$
\end{itemize}
\end{enumerate}
\end{ex}

\begin{rem}
Note that if $\Gamma$ is the graph associated to a Dynkin diagram, then $W_\Gamma$ as we have defined it is precisely the Artin group corresponding to that Dynkin diagram. This occurs because, in this case, we have no cycles in $\Gamma,$ and so we only have relation of the form $(T2'),$ which define the Artin group.
\end{rem}
